---
title: "CS 422"
author: "Nikolaus Schultze, Illinois Institute of Technology"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
---

<!-- More information in R Markdown can be found at:
1. https://www.ssc.wisc.edu/sscc/pubs/RFR/RFR_RMarkdown.html  This is 
   the place to start since it is a short tutorial.
2. https://rmarkdown.rstudio.com/index.html This contains a longer 
   tutorial.  Take a look at the cheatsheet in 
   https://rmarkdown.rstudio.com/lesson-15.html, it is a concise 
   reference of R Markdown on two pages.
<-->

## Use this as a template for your homeworks.
#### Rename it to firstname-lastname.Rmd.
#### Run all the chunks by clicking on "Run" at the top right of the edit 
#### window and choose "Run All".  Assuming there were no errors in the
#### chunk, you should see a "Preview" button become visible on the top
#### left of the edit window.  Click this button and a html document should
#### pop up with the output from this R markdown script.

### Part 2.1-A
```{r}
hotel_bookings = read.csv(file = '.\\hotel_bookings.csv')
setwd("~/RTut")
#hotel_bookings
hotel_type = data.frame(table(hotel_bookings$hotel))
#hotel_type
print(paste0("There are ", hotel_type[2,2], " hotel type H1s"))
print(paste0("There are ", hotel_type[1,2], " hotel type H2s"))
```

### Part 2.1-B.

```{r}
cancel = data.frame(table(hotel_bookings$is_canceled))
#cancel
print(paste0("Number of guests who canceled reservation: ", cancel[1,2]))
print(paste0("Number of guests who did not cancel the reservation: ", cancel[2,2]))
```
### Part 2.1-C
```{r}
customer_type = data.frame(table(hotel_bookings$customer_type))
#customer_type
max_customer_type = which.max(customer_type$Freq)
#max_customer_type
print(paste0("Customer type with the most reservations is ", customer_type[max_customer_type, 1], ", with ", customer_type[max_customer_type, 2]," reservations"))
```

### Part 2.1-D
```{r}
parking_spaces = data.frame(table(hotel_bookings$required_car_parking_spaces))
#parking_spaces
max_parking_spaces = which.max(parking_spaces$Var1)
#max_parking_spaces
print(paste0(parking_spaces[max_parking_spaces, 2]," customers required the most number of parking spaces (", parking_spaces[max_parking_spaces, 1], ")"))
```

### Part 2.1-E
```{r}
min_parking_spaces = which.min(parking_spaces$Var1)
#min_parking_spaces
print(paste0(parking_spaces[min_parking_spaces, 2]," customers required the least number of parking spaces (", parking_spaces[min_parking_spaces, 1], ")"))
```

### Part 2.1-F
```{r}
room_type = hotel_bookings[,c(20:21)]
#room_type
new_column_room_type = data.frame(table(ifelse(room_type$reserved_room_type == room_type$assigned_room_type, '1', '0')))
#new_column_room_type
room_preference_percentage = ((new_column_room_type[2,2]) / (new_column_room_type[1,2] + new_column_room_type[2,2])) * 100
#room_preference_percentage
print(paste0(round(room_preference_percentage, digits = 2),"% of the people who expressed a room preference during reservation got the room during check-in."))
```

### Part 2.1-G
```{r}
#install.packages('tidyverse')
library(tidyverse)
resort_hotel_bookings = hotel_bookings[hotel_bookings$hotel == "Resort Hotel",]
city_hotel_bookings = hotel_bookings[hotel_bookings$hotel == "City Hotel",]

resort_hotel_bookings_sorted = table(resort_hotel_bookings$country)
resort_hotel_bookings_sorted_order = sort(resort_hotel_bookings_sorted, decreasing = TRUE)

city_hotel_bookings_sorted = table(city_hotel_bookings$country)
city_hotel_bookings_sorted_order = sort(city_hotel_bookings_sorted, decreasing = TRUE)

new_resort_hotel_bookings_sorted_order = resort_hotel_bookings_sorted_order[c(1:9,11)]
new_city_hotel_bookings_sorted_order = city_hotel_bookings_sorted_order[c(1:10)]

new_resort_hotel_bookings_sorted_order = sort(new_resort_hotel_bookings_sorted_order, decreasing = FALSE)
new_city_hotel_bookings_sorted_order = sort(new_city_hotel_bookings_sorted_order, decreasing = FALSE)

barplot(new_resort_hotel_bookings_sorted_order, main = "Top 10 Countries of Origin for Resort Hotels",
        col = rainbow(7), xlab = "Country")
barplot(new_city_hotel_bookings_sorted_order, main = "Top 10 Countries of Origin for City Hotels",
        col = rainbow(11), xlab = "Country")
```

### Part 2.1-H(i)
Portugal

### Part 2.1-H(ii)
The Origin of the dataset shows the a large portion of both Resort and City Hotels are located in Portugal.

### Part 2.2-(A)(i)
```{r}
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("psych")
library(rpart)
library(rpart.plot)
library(psych)
#install.packages("caret")
library(caret)
#install.packages("caTools")
library(caTools)
set.seed(1122)
sample = sample(1:nrow(hotel_bookings), .90 * dim(hotel_bookings)[1])
train = hotel_bookings[sample,]
test = hotel_bookings[-sample,]
#train
#test

train_correlation1 = train[,c(2, 1, 3:10)]
train_correlation2 = train[,c(2, 11:21)]
train_correlation3 = train[,c(2, 22, 31)]

#pairs.panels(train_correlation1)
#pairs.panels(train_correlation2)
#pairs.panels(train_correlation3)
decision_tree = rpart(is_canceled ~ deposit_type + lead_time + previous_cancellations + market_segment + total_of_special_requests + required_car_parking_spaces, data = train, method = 'class')
#decision_tree
rpart.plot(decision_tree, extra=104, fallen.leaves = T, type = 4, main = "Decision Tree")
```

### Part 2.2-A(ii)
The variables that are important are Deposit Type, Lead Time, Previous Cancellations, Market Segment, Total of Special Requests, and Required Car Parking Spaces.

### Part 2.2-A(iii)
```{r}
decision_tree_predicted = predict(decision_tree, test, type = 'class')
confusion_matrix = confusionMatrix(decision_tree_predicted, as.factor(test$is_canceled))
confusion_matrix

accuracy = round(confusion_matrix$overall['Accuracy'], digits = 3)
error = round(1- accuracy, digits = 3)
balanced_accuracy = round(confusion_matrix$byClass['Balanced Accuracy'], digits = 3)
specificity = round(confusion_matrix$byClass['Specificity'], digits = 3)
sensitivity = round(confusion_matrix$byClass['Sensitivity'], digits = 3)
precision = round(confusion_matrix$byClass['Pos Pred Value'], digits = 3)

print(paste0("The Accuracy is : ", + accuracy))
print(paste0("The Error is : ", + error))
print(paste0("The Balanced Accuracy is : ", + balanced_accuracy))
print(paste0("The Specificity is : ", + specificity))
print(paste0("The Sensitivity is : ", + sensitivity))
print(paste0("The Precision is : ", + precision))
```

### Part 2.2-A(iv)
```{r}
#install.packages("ROCR")
library("pROC")
roc_curve = roc(response = test$is_canceled, predictor = factor(decision_tree_predicted, order = TRUE), auc = TRUE)
plot(roc_curve, print.auc = TRUE)
```

### Part 2.2-A(v)
The AUC of the ROC curve is 0.766.

### Part 2.3-A
```{r}
set.seed(1122)
sample = sample(1:nrow(hotel_bookings), .90 * dim(hotel_bookings)[1])
train2 = hotel_bookings[sample,]
test2 = hotel_bookings[-sample,]
#train2
#test2

decision_tree_cp_0 = rpart(is_canceled ~ deposit_type + lead_time + previous_cancellations + market_segment + total_of_special_requests + required_car_parking_spaces, data = train2, method = 'class', cp = 0)
predicted_decision_tree_cp_0 = predict(decision_tree_cp_0, test2, type = 'class')
confusion_matrix = confusionMatrix(predicted_decision_tree_cp_0, factor(test$is_canceled))
confusion_matrix

prune_accuracy = round(confusion_matrix$overall['Accuracy'], digits = 3)
prune_error = round(1- accuracy, digits = 3)
prune_balanced_accuracy = round(confusion_matrix$byClass['Balanced Accuracy'], digits = 3)
prune_specificity = round(confusion_matrix$byClass['Specificity'], digits = 3)
prune_sensitivity = round(confusion_matrix$byClass['Sensitivity'], digits = 3)
prune_precision = round(confusion_matrix$byClass['Pos Pred Value'], digits = 3)

print(paste0("Before Pruning:"))
print(paste0("Accuracy: ", + prune_accuracy))
print(paste0("Error: ", + prune_error))
print(paste0("Balanced Acc.: ", + prune_balanced_accuracy))
print(paste0("Specificity: ", + prune_specificity))
print(paste0("Sensitivity: ", + prune_sensitivity))
print(paste0("Precision: ", prune_precision))
```

### Part 2.3-B
```{r}
#printcp(decision_tree)
#plotcp(decision_tree)
cpx = decision_tree$cptable[which.min(decision_tree$cptable[, "xerror"]), "CP"]
pruned.model = prune(decision_tree_cp_0, cp = 0.325791)
pruned.model
pred = predict(pruned.model, test, type = "class")
pruned_confusion_matrix = confusionMatrix(pred, as.factor(test2$is_canceled))
pruned_confusion_matrix
print(paste0("Prune point occurs at a complexity of 0.325791"))
print(paste0("At this complexity, xerror is 1.00"))

```

### Part 2.3-C
```{r}
pruned_accuracy = round(pruned_confusion_matrix$overall['Accuracy'], digits = 3)
pruned_error = round(1- accuracy, digits = 3)
pruned_balanced_accuracy = round(pruned_confusion_matrix$byClass['Balanced Accuracy'], digits = 3)
pruned_specificity = round(pruned_confusion_matrix$byClass['Specificity'], digits = 3)
pruned_sensitivity = round(pruned_confusion_matrix$byClass['Sensitivity'], digits = 3)
pruned_precision = round(pruned_confusion_matrix$byClass['Pos Pred Value'], digits = 3)

print(paste0("After Pruning:"))
print(paste0("Accuracy: ", + pruned_accuracy))
print(paste0("Error: ", + pruned_error))
print(paste0("Balanced Acc.: ", + pruned_balanced_accuracy))
print(paste0("Specificity: ", + pruned_specificity))
print(paste0("Sensitivity: ", + pruned_sensitivity))
print(paste0("Precision: ", + pruned_precision))
```

### Part 2.3-D
The Full tree generalizes better.

### Part 2.3-E
The model in Problem 2.3(A) generalizes the best.
