---
title: "CS 422"
author: "Nikolaus Schultze, Illinois Institute of Technology"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
---

<!-- More information in R Markdown can be found at:
1. https://www.ssc.wisc.edu/sscc/pubs/RFR/RFR_RMarkdown.html  This is 
   the place to start since it is a short tutorial.
2. https://rmarkdown.rstudio.com/index.html This contains a longer 
   tutorial.  Take a look at the cheatsheet in 
   https://rmarkdown.rstudio.com/lesson-15.html, it is a concise 
   reference of R Markdown on two pages.
<-->

## Use this as a template for your homeworks.
#### Rename it to firstname-lastname.Rmd.
#### Run all the chunks by clicking on "Run" at the top right of the edit 
#### window and choose "Run All".  Assuming there were no errors in the
#### chunk, you should see a "Preview" button become visible on the top
#### left of the edit window.  Click this button and a html document should
#### pop up with the output from this R markdown script.
install.packages("dplyr")
install.packages("psych")

### Part 2.1-A
```{r}
college = read.csv(file = '.\\College.csv')
setwd("~/RTut")
college[c(1:5), c(1, 5, 8, 10)]
```

### Part 2.1-B
```{r}
private = sum(college[2] == "Yes")
public = sum(college[2] == "No")

str = paste0("There are ", private, " private colleges and ", public, " public colleges in the datatset")
str
```
### Part 2.1-C 
```{r}
dataframe = college[,c(2, 3, 4, 5, 14, 17, 16, 19)]
dataframe[1:6,]
```

### Part 2.1-D(i)
```{r}
private_yes = dataframe$Private == 'Yes'
private_college = dataframe[private_yes,]
#private_college
hist(private_college[,5], col = rainbow(2))
```

### Part 2.1-D(ii)
```{r}
private_no = dataframe$Private == "No"
public_college = dataframe[private_no,]
#public_college
Color = rainbow(54)
hist(public_college[,5], col = rainbow(9))
```

### Part 2.1-E(i)
```{r}
min_grad_rate = dataframe[order(dataframe[,8]),]
min_grad_rate[c(1:5), c(8)]
```

### Part 2.1-E(ii)
```{r}
max_grad_rate = dataframe[order(dataframe[,8], decreasing = TRUE),]
max_grad_rate[c(1:6), c(8)]
```

### Part 2.1-F(i)
```{r}
library(psych)
dataframe_cor = dataframe[,c(5, 7, 8)]
pairs.panels(dataframe_cor)
```

### Part 2.1-F(ii)
```{r}
print("PhD and Grad Rate have the highest correlation. This makes sense because teachers with PhDs might bew considered by some to be better teacheres teachers therefore they improve graduation rates.")
```

### Part 2.1-F(iii)
```{r}
print("S.F Ratio and Grad Rate have the lowest correlation. This makes sense because a good Student to Faculty ratio is lower meaning there is less students per faculty member. This would allow these faculty members to spend more time and focus on students meaning that the students would do better and thus Graduation rates would be higher.")
```

### Part 2.1-G
```{r}
boxplot(college$perc.alumni~(college$Private == 'Yes'), xlab = "Public College and Private College", ylab = "Percen of Alumni who Donate", main = "Private vs Public Alumni Donation Percentage")
print("Alumni that went to Private colleges tend to donate more money that Alumni that went to Public colleges.")
```
### Part 2.1-H(iii)
```{r}
ecdf(college$Expend)
plot(ecdf(college$Expend))
grid(10000, 10000)
```

### Part 2.1-H(i)
```{r}
print("The Median Expenditure per Student is $8,500")
```

### Part 2.1-H(ii)
```{r}
print(" 80% of the students pay less than $10,000")
```
