---
title: "CS 422"
author: "Nikolaus Schultze, Illinois Institute of Technology"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
---

<!-- More information in R Markdown can be found at:
1. https://www.ssc.wisc.edu/sscc/pubs/RFR/RFR_RMarkdown.html  This is 
   the place to start since it is a short tutorial.
2. https://rmarkdown.rstudio.com/index.html This contains a longer 
   tutorial.  Take a look at the cheatsheet in 
   https://rmarkdown.rstudio.com/lesson-15.html, it is a concise 
   reference of R Markdown on two pages.
<-->

## Use this as a template for your homeworks.
#### Rename it to firstname-lastname.Rmd.
#### Run all the chunks by clicking on "Run" at the top right of the edit 
#### window and choose "Run All".  Assuming there were no errors in the
#### chunk, you should see a "Preview" button become visible on the top
#### left of the edit window.  Click this button and a html document should
#### pop up with the output from this R markdown script.

### Part 2.1-A
```{r}
tr1ks_canonical = read.csv(file = '.\\tr-1k-canonical.csv', sep = '\t', header = FALSE)
tr5ks_canonical = read.csv(file = '.\\tr-5k-canonical.csv', sep = '\t', header = FALSE)
tr20ks_canonical = read.csv(file = '.\\tr-20k-canonical.csv', sep = '\t', header = FALSE)
tr75ks_canonical = read.csv(file = '.\\tr-75k-canonical.csv', sep = '\t', header = FALSE)
```

### Part 2.1-B
```{r}
#install.packages("arules")
#install.packages("arulesViz")
library(arules)
library(arulesViz)
trans <- read.transactions("tr-1k-canonical.csv", sep=",")
f_is <- apriori(trans, parameter=list(support=0.09, target="frequent itemsets"))
inspect(sort(f_is, decreasing = T, by="count"))
rm(f_is)
rules <- apriori(trans, parameter = list(support=0.01))
summary(rules)


trans5k <- read.transactions("tr-5k-canonical.csv", sep=",")
f_is5k <- apriori(trans5k, parameter=list(support=0.09, target="frequent itemsets"))
inspect(sort(f_is5k, decreasing = T, by="count"))
rm(f_is5k)
rules5k <- apriori(trans5k, parameter = list(support=0.01))
summary(rules5k)

trans20k <- read.transactions("tr-20k-canonical.csv", sep=",")
f_is20k <- apriori(trans20k, parameter=list(support=0.09, target="frequent itemsets"))
inspect(sort(f_is20k, decreasing = T, by="count"))
rm(f_is20k)
rules20k <- apriori(trans20k, parameter = list(support=0.01))
summary(rules20k)

trans75k <- read.transactions("tr-75k-canonical.csv", sep=",")
f_is75k <- apriori(trans75k, parameter=list(support=0.09, target="frequent itemsets"))
inspect(sort(f_is75k, decreasing = T, by="count"))
rm(f_is75k)
rules75k <- apriori(trans75k, parameter = list(support=0.01))
summary(rules75k)
```

### Part 2.1-C
The number of transactions affect the results as the 1k had a set of 84 rules, the 5k had a set of 84 rules, the 20k had a set of 84 rules, but the 75k had a set of 85 rules.


### Part 2.1-D(i)
```{r}
print("The most frequently purchased item or itemset is the Coffee Eclair")
```

### Part 2.1-D(ii)
```{r}
print("The least frequently purchased item or itemset is the Gongolais Cookie")
```

### Part 2.2
```{r}
library(tidyverse)
#install.packages("lsa")
library(lsa)
#20442006 mod 671 = 662
ratings = read.csv(file = '.\\ratings.csv')
movies = read.csv(file = '.\\movies.csv')
#ratings
#movies
randomMovies = sample(1:9125, 10)
#randomMovies
recommendedMovies = movies[c(randomMovies), c(1,2)]

user = 1
userMovies = ratings %>% filter(userId == user)
userMoviesID = data.frame(userMovies[,c(2)])
#userMoviesID
numberOfMovies = length(userMovies$movieId)
#numberOfMovies
#userProfile = data.frame(userMovies$movieId)

newMovies = data.frame(matrix(ncol = 3, nrow = 0))
colnames(newMovies) = c("movieId", "title", "genres")
i= 1
repeat {
  newMoviesTemp = movies %>% filter(movieId == userMoviesID[i,])
  paste(newMoviesTemp)
  newMovies = rbind(newMovies, newMoviesTemp)
  i = i + 1
  if(i == length(userMovies$movieId) + 1) {
    break
  }
}
#newMovies

userProfile = data.frame(matrix(ncol = 21, nrow = 0))
colnames(userProfile) = c("Movie ID", "Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy", "Film-Noir", "Horror", "IMAX", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western", "(no genres listed)")
userProfileFinal = data.frame(matrix(ncol = 21, nrow = 0))
colnames(userProfileFinal) = c("Movie ID", "Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy", "Film-Noir", "Horror", "IMAX", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western", "(no genres listed)")
#userProfile
j = 1
repeat {
  if(grepl("Action", newMovies$genres[j]))
    (userProfile[1,1] = newMovies$movieId[j]) && (userProfile[1,2] = 1)
  else
    (userProfile[1,1] = newMovies$movieId[j]) && (userProfile[1,2] = 0)
  if(grepl("Adventure", newMovies$genres[j]))
    userProfile[1,3] = 1
  else
    userProfile[1,3] = 0
  if(grepl("Animation", newMovies$genres[j]))
    userProfile[1,4] = 1
  else
    userProfile[1,4] = 0
  if(grepl("Children", newMovies$genres[j]))
    userProfile[1,5] = 1
  else
    userProfile[1,5] = 0
  if(grepl("Comedy", newMovies$genres[j]))
    userProfile[1,6] = 1
  else
    userProfile[1,6] = 0
  if(grepl("Crime", newMovies$genres[j]))
    userProfile[1,7] = 1
  else
    userProfile[1,7] = 0
  if(grepl("Documentary", newMovies$genres[j]))
    userProfile[1,8] = 1
  else
    userProfile[1,8] = 0
  if(grepl("Drama", newMovies$genres[j]))
    userProfile[1,9] = 1
  else
    userProfile[1,9] = 0
  if(grepl("Fantasy", newMovies$genres[j]))
    userProfile[1,10] = 1
  else
    userProfile[1,10] = 0
  if(grepl("Film-Noir", newMovies$genres[j]))
    userProfile[1,11] = 1
  else
    userProfile[1,11] = 0
  if(grepl("Horror", newMovies$genres[j]))
    userProfile[1,12] = 1
  else
    userProfile[1,12] = 0
  if(grepl("IMAX", newMovies$genres[j]))
    userProfile[1,13] = 1
  else
    userProfile[1,13] = 0
  if(grepl("Musical", newMovies$genres[j]))
    userProfile[1,14] = 1
  else
    userProfile[1,14] = 0
  if(grepl("Mystery", newMovies$genres[j]))
    userProfile[1,15] = 1
  else
    userProfile[1,15] = 0
  if(grepl("Romance", newMovies$genres[j]))
    userProfile[1,16] = 1
  else
    userProfile[1,16] = 0
  if(grepl("Sci-Fi", newMovies$genres[j]))
    userProfile[1,17] = 1
  else
    userProfile[1,17] = 0
  if(grepl("Thriller", newMovies$genres[j]))
    userProfile[1,18] = 1
  else
    userProfile[1,18] = 0
  if(grepl("War", newMovies$genres[j]))
    userProfile[1,19] = 1
  else
    userProfile[1,19] = 0
  if(grepl("Western", newMovies$genres[j]))
    userProfile[1,20] = 1
  else
    userProfile[1,20] = 0
  if(grepl("(no genres listed)", newMovies$genres[j]))
    userProfile[1,21] = 1
  else
    userProfile[1,21] = 0
  userProfileFinal = rbind(userProfileFinal, userProfile)
  j = j + 1
  if(j == length(userMovies$movieId) + 1) {
    break
  }
}
#userProfileFinal

avgAction = sum(userProfileFinal$Action) / length(userMovies$movieId)
avgAdventure = sum(userProfileFinal$Adventure) / length(userMovies$movieId)
avgAnimation = sum(userProfileFinal$Animation) / length(userMovies$movieId)
avgChildren = sum(userProfileFinal$Children) / length(userMovies$movieId)
avgComedy = sum(userProfileFinal$Comedy) / length(userMovies$movieId)
avgCrime = sum(userProfileFinal$Crime) / length(userMovies$movieId)
avgDocumentary = sum(userProfileFinal$Documentary) / length(userMovies$movieId)
avgDrama = sum(userProfileFinal$Drama) / length(userMovies$movieId)
avgFantasy = sum(userProfileFinal$Fantasy) / length(userMovies$movieId)
avgFilmNoir = sum(userProfileFinal$`Film-Noir`) / length(userMovies$movieId)
avgHorror = sum(userProfileFinal$Horror) / length(userMovies$movieId)
avgIMAX = sum(userProfileFinal$IMAX) / length(userMovies$movieId)
avgMusical = sum(userProfileFinal$Musical) / length(userMovies$movieId)
avgMystery = sum(userProfileFinal$Mystery) / length(userMovies$movieId)
avgRomance = sum(userProfileFinal$Romance) / length(userMovies$movieId)
avgSciFi = sum(userProfileFinal$`Sci-Fi`) / length(userMovies$movieId)
avgThriller = sum(userProfileFinal$Thriller) / length(userMovies$movieId)
avgWar = sum(userProfileFinal$War) / length(userMovies$movieId)
avgWestern = sum(userProfileFinal$Western) / length(userMovies$movieId)
avgNoGenre = sum(userProfileFinal$`(no genres listed)`) / length(userMovies$movieId)
avg = c("AVG", avgAction, avgAdventure, avgAnimation, avgChildren, avgComedy, avgCrime, avgDocumentary, avgDrama, avgFantasy, avgFilmNoir, avgHorror, avgIMAX, avgMusical, avgMystery, avgRomance, avgSciFi, avgThriller, avgWar, avgWestern, avgNoGenre)
#avg

userProfileVector = c(avgAction, avgAdventure, avgAnimation, avgChildren, avgComedy, avgCrime, avgDocumentary, avgDrama, avgFantasy, avgFilmNoir, avgHorror, avgIMAX, avgMusical, avgMystery, avgRomance, avgSciFi, avgThriller, avgWar, avgWestern, avgNoGenre)

userProfileFinal = rbind(userProfileFinal, avg)
#userProfileFinal

movieProfile = data.frame(matrix(ncol = 20, nrow = 0))
colnames(movieProfile) = c("Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy", "Film-Noir", "Horror", "IMAX", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western", "(no genres listed)")
movieProfileFinal = data.frame(matrix(ncol = 20, nrow = 0))
colnames(movieProfileFinal) = c("Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy", "Film-Noir", "Horror", "IMAX", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western", "(no genres listed)")
k = randomMovies
j = 1
repeat {
  if(grepl("Action", movies$genres[k[j]]))
    movieProfile[1,1] = 1
  else
    movieProfile[1,1] = 0
  if(grepl("Adventure", movies$genres[k[j]]))
    movieProfile[1,2] = 1
  else
    movieProfile[1,2] = 0
  if(grepl("Animation", movies$genres[k[j]]))
    movieProfile[1,3] = 1
  else
    movieProfile[1,3] = 0
  if(grepl("Children", movies$genres[k[j]]))
    movieProfile[1,4] = 1
  else
    movieProfile[1,4] = 0
  if(grepl("Comedy", movies$genres[k[j]]))
    movieProfile[1,5] = 1
  else
    movieProfile[1,5] = 0
  if(grepl("Crime", movies$genres[k[j]]))
    movieProfile[1,6] = 1
  else
    movieProfile[1,6] = 0
  if(grepl("Documentary", movies$genres[k[j]]))
    movieProfile[1,7] = 1
  else
    movieProfile[1,7] = 0
  if(grepl("Drama", movies$genres[k[j]]))
    movieProfile[1,8] = 1
  else
    movieProfile[1,8] = 0
  if(grepl("Fantasy", movies$genres[k[j]]))
    movieProfile[1,9] = 1
  else
    movieProfile[1,9] = 0
  if(grepl("Film-Noir", movies$genres[k[j]]))
    movieProfile[1,10] = 1
  else
    movieProfile[1,10] = 0
  if(grepl("Horror", movies$genres[k[j]]))
    movieProfile[1,11] = 1
  else
    movieProfile[1,11] = 0
  if(grepl("IMAX", movies$genres[k[j]]))
    movieProfile[1,12] = 1
  else
    movieProfile[1,12] = 0
  if(grepl("Musical", movies$genres[k[j]]))
    movieProfile[1,13] = 1
  else
    movieProfile[1,13] = 0
  if(grepl("Mystery", movies$genres[k[j]]))
    movieProfile[1,14] = 1
  else
    movieProfile[1,14] = 0
  if(grepl("Romance", movies$genres[k[j]]))
    movieProfile[1,15] = 1
  else
    movieProfile[1,15] = 0
  if(grepl("Sci-Fi", movies$genres[k[j]]))
    movieProfile[1,16] = 1
  else
    movieProfile[1,16] = 0
  if(grepl("Thriller", movies$genres[k[j]]))
    movieProfile[1,17] = 1
  else
    movieProfile[1,17] = 0
  if(grepl("War", movies$genres[k[j]]))
    movieProfile[1,18] = 1
  else
    movieProfile[1,18] = 0
  if(grepl("Western", movies$genres[k[j]]))
    movieProfile[1,19] = 1
  else
    movieProfile[1,19] = 0
  if(grepl("(no genres listed)", movies$genres[k[j]]))
    movieProfile[1,20] = 1
  else
    movieProfile[1,20] = 0
  movieProfileFinal = rbind(movieProfileFinal, movieProfile)
  j = j + 1
  if(j == 10) {
    break
  }
}
#movieProfileFinal

movieOne =  as.vector(t(movieProfileFinal[1,]))
movieTwo = as.vector(t(movieProfileFinal[2,]))
movieThree = as.vector(t(movieProfileFinal[3,]))
movieFour = as.vector(t(movieProfileFinal[4,]))
movieFive = as.vector(t(movieProfileFinal[5,]))
movieSix = as.vector(t(movieProfileFinal[6,]))
movieSeven = as.vector(t(movieProfileFinal[7,]))
movieEight = as.vector(t(movieProfileFinal[8,]))
movieNine = as.vector(t(movieProfileFinal[9,]))
movieTen = as.vector(t(movieProfileFinal[10,]))

sim1 = round(cosine(userProfileVector, movieOne), digits = 3)
sim2 = round(cosine(userProfileVector, movieTwo), digits = 3)
sim3 = round(cosine(userProfileVector, movieThree), digits = 3)
sim4 = round(cosine(userProfileVector, movieFour), digits = 3)
sim5 = round(cosine(userProfileVector, movieFive), digits = 3)
sim6 = round(cosine(userProfileVector, movieSix), digits = 3)
sim7 = round(cosine(userProfileVector, movieSeven), digits = 3)
sim8 = round(cosine(userProfileVector, movieEight), digits = 3)
sim9 = round(cosine(userProfileVector, movieNine), digits = 3)
sim10 = round(cosine(userProfileVector, movieTen), digits = 3)

recommendedMovies$Similarity = c(sim1, sim2, sim3, sim4, sim5, sim6, sim7, sim8, sim9, sim10)
recommendedMovies = recommendedMovies[order(-recommendedMovies$Similarity),]
recommendedMoviesFinal = recommendedMovies[c(1:5),]

print("User 662 chose the following 10 movies: ")
paste0(randomMovies, sep = ',', collapse = ' ')
print("Of these, the following 5 movies are recommended:")
recommendedMoviesFinal
```