---
title: "CS 422"
author: "Nikolaus Schultze, Illinois Institute of Technology"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
---

<!-- More information in R Markdown can be found at:
1. https://www.ssc.wisc.edu/sscc/pubs/RFR/RFR_RMarkdown.html  This is 
   the place to start since it is a short tutorial.
2. https://rmarkdown.rstudio.com/index.html This contains a longer 
   tutorial.  Take a look at the cheatsheet in 
   https://rmarkdown.rstudio.com/lesson-15.html, it is a concise 
   reference of R Markdown on two pages.
<-->

## Use this as a template for your homeworks.
#### Rename it to firstname-lastname.Rmd.
#### Run all the chunks by clicking on "Run" at the top right of the edit 
#### window and choose "Run All".  Assuming there were no errors in the
#### chunk, you should see a "Preview" button become visible on the top
#### left of the edit window.  Click this button and a html document should
#### pop up with the output from this R markdown script.

### Part 1-A
```{r}
#setwd("/tmp")
data(cars)
cars
attach(cars)
```

### Part 1-B.
```{r}
plot(speed, dist, main = "Cars", xlab = "Speed", ylab = "Distance") 
plot(cars, type = "o")
```
### Part 1-C (Summary)
```{r}
print(summary(cars))
```
### Part 1-D
```{r}
# Add code here.  Save maximum speed in a variable called "max.speed" and
# minimum distance in a variable called "min.dist".  The code below will
# print your output.
max.speed = max(speed)
min.dist  = min(dist)
str <- paste0("Maximum speed is ", max.speed, ", and minimum distance is ", 
              min.dist)
str
```
